---
title: "Lab 10"
author: "Nikita Grabher-Meyer"
date: "11/11/2020"
output:
  word_document: default
  html_document: default
---

# Matrix, Figures and Experiments

## 1) Starting from file Lab_09.RScript, try to run yourself what you saw in "Matrix_03" the primer session

*Identity matrix*
```{r}
I1 <- diag(1)
I1

I2 <- diag(2)
I2

A <- matrix(1:4, nrow=2, ncol = 2)
A

AI <- A%*%I2
AI
```

*Symmetric matrix*
```{r}
S <- cbind(c(1,2,3),c(2,1,4), c(3,4,3))
S

trS <- t(S)
trS
```

*Basic properties*
```{r}
X<- cbind(c(1,2), c(3,4), c(5,6))
Y <- cbind(c(7,8), c(9,10), c(11,12))
X+Y

Z <- cbind(c(1,2), c(3,4))
W <- cbind(c(5,6), c(8,8))
ZW1 <- Z%*%W
ZW2 <- W%*%Z
ZW1
ZW2

trZW1 <- t(ZW1)
trZ <- t(Z)
trW <- t(W)
trWtrZ <- trW%*%trZ
trZW1
trWtrZ

trtrZ <- t(trZ)
trtrZ
Z
```

*Inverse matrix*
```{r}
C <- matrix(c(1,0,3,1), nrow=2, ncol = 2)
C

invC <- solve(C)
invC

CinvC <- C%*%invC
CinvC
```


## 2) Using the file Lab02_mkm_Econometrics1_RodBased_DataHandling: 

### a.Read dt.wages and apply ‘data.table’
```{r}
setwd("/Users/nikitagrabher-meyer/Desktop/PHD/Econometrics/Labs/Lab 10, Homework")

library(data.table)
library(ggplot2)
library(stargazer)
library(GGally)
library(Metrics)

load("dt_wages.RData")
```

```{r}
dt.wages <- data.table(dt.wages)
```

### b.Get basic summary statistics. How many observations, how many variables?
```{r}
str(dt.wages)
head(dt.wages, 3)
tail(dt.wages, 3)
summary(dt.wages)
stargazer (dt.wages, type="text")
stargazer (dt.wages, type="text", iqr=TRUE)
ncol(dt.wages)

table(dt.wages[, list(female, nonwhite)])
table(dt.wages[, list (female, nonwhite, south)])
table(dt.wages[, list (female , tenure)])
table(dt.wages[, list (nonwhite , tenure)])
```

## c.Reproduce the data-handling steps shown in the slides (and dt.wages)
```{r}
qplot(factor(educ),data=dt.wages,geom="bar")
qplot(x=factor(1),fill=factor(educ),data=dt.wages,geom="bar") + coord_polar(theta="y")
qplot(wage,data=dt.wages,geom="histogram")
qplot(wage,data=dt.wages,geom="histogram",binwidth=2)
qplot(x=factor("All"),y=wage,data=dt.wages,geom="boxplot")
qplot(factor(female),wage,data=dt.wages,geom="boxplot")
qplot(factor(female),wage,fill=factor(nonwhite),data=dt.wages,geom="boxplot")
qplot(factor(nonwhite),wage,fill=factor(female),data=dt.wages,geom="boxplot")
ggplot(dt.wages)+geom_boxplot(aes(factor(educ),exper))
qplot(educ,wage,data=dt.wages,geom="point")
qplot(educ,wage,color=factor(female),data=dt.wages,geom="point")
ggplot(dt.wages)+geom_point(aes(educ,wage))+facet_grid(~female)
ggplot(dt.wages)+geom_point(aes(educ,wage))+facet_grid(nonwhite~female)
ggplot(dt.wages)+geom_point(aes(educ,wage))+facet_grid(nonwhite~female, margins=TRUE)
ggpairs(dt.wages[,list(wage,educ,exper)])
ggpairs(dt.wages[,list(female=factor(female),nonwhite=factor(nonwhite))])
ggpairs(dt.wages[,list(wage,educ,exper,female=factor(female),nonwhite=factor(nonwhite))])
```

```{r}
dt.wages[wage>20,][order(-wage)]
qplot(lwage,data=dt.wages,geom="histogram")
dt.wages[lwage< 0,]
```


## 3) Let’s do a thought experiment: Using the slide-deck UEA_ecoR2PhD CoreLect_06 ATENT_Match_Stk and dt.wages, let’s pretend, for a moment “south,” is a randomly assigned treatment in an experiment:

### a.Compute a difference-in-means estimator when treatment is “south,” and the outcome is wage
```{r}
dt.wages[south==1, mean(wage)] - dt.wages[south==0, mean(wage)]
```

### b.Now focus on race and gender as control variables (in “x”) and run a regression estimation of treatment effects
```{r}
reg1 <- lm(wage ~ south + nonwhite + female, data=dt.wages)
stargazer(reg1, type = "text")
```

### c.Now try to estimate the regression and account for potentially heterogeneous treatment effects. 
```{r}
reg2 <- lm(wage ~ south + nonwhite + female + nonwhite*south + female*south, data=dt.wages)
stargazer(reg2, type = "text")
```

### d.Try to implement a 2-step fitted regression.
```{r}
south.yes <- subset(dt.wages, south==1)
south.no <- subset(dt.wages, south==0)

reg3.1 <- lm(wage ~ nonwhite + female, data=south.yes)
reg3.2 <- lm(wage ~ nonwhite + female, data=south.no)

stargazer(reg3.1, reg3.2, type = "text")

dt.wages$y_hat1 <- predict(reg3.1, newdata=dt.wages)
dt.wages$y_hat0 <- predict(reg3.2, newdata=dt.wages)
```
