---
title: "Homework 2"
author: "Nikita Grabher-Meyer"
date: "10/27/2020"
output: html_document
---


## Hlavac, Marek (2018). stargazer: Well-Formatted Regression and Summary Statistics Tables.
## R package version 5.2.2. https://CRAN.R-project.org/package=stargazer


### Setup: set working directory, install and activate necessary packages

```{r setup, include=FALSE}
setwd("/Users/nikitagrabher-meyer/Desktop/PHD/Econometrics")
library("ggplot2")
library("stargazer")
library("data.table")
```

# Data Analysis

### Load data set
```{r} 
load("/Users/nikitagrabher-meyer/Desktop/PHD/Econometrics/ceosal2.RData")
dt.ceo.salaries <- data.table(data)
rm(data)
```

### Descriptive statistics

*How many CEOS are in the sample?*
```{r} 
nrow(dt.ceo.salaries)
```

*How many CEOS have a graduate degree?*
```{r} 
nrow(dt.ceo.salaries[grad==1,])
```

*What is the percentage of CEOs with graduate degrees?*
```{r} 
dt.ceo.salaries[, sum(grad)]/nrow(dt.ceo.salaries)
```

*What is the average CEO salary?*
```{r} 
dt.ceo.salaries[, mean(salary)]
```

*What is the mean CEO salary for those with a graduate degree?*
```{r} 
dt.ceo.salaries[grad==1, mean(salary)]
```

*What is the mean CEO salary for those without a graduate degree?*
```{r} 
dt.ceo.salaries[grad==0, mean(salary)]
```

*How many CEOs have/don’t have a college degree?*
```{r}
dt.ceo.salaries[ , list(n_ceo=.N), by = college]
```

*Can we say that the mean salary is statistically different from 800?*
```{r} 
t.test(dt.ceo.salaries[, salary], mu = 800)
```

We can not reject the null Hp that the population mean is = 800 (this value is indeed contained in the 95 percent confidence interval: 778.7015 953.0274)


*Is the average salary different for CEOs with a graduate degree and those without?*
```{r} 
t.test(dt.ceo.salaries[grad==0, salary] , dt.ceo.salaries[grad==1, salary])
```

We can not reject the null Hp that there is no difference between the two population means (the value 0 - no difference between the two means - is indeed contained in the 95 percent confidence interval: -175.0489 182.0932)


*Summary table of descriptive statistics*
```{r} 
stargazer(dt.ceo.salaries, type = "text")
```


### Plots

**Histogram**

*Salary*
```{r pressure, echo=FALSE}
qplot( data = dt.ceo.salaries
, x = salary
, geom = "histogram")
```

*Age*
```{r}
qplot( data = dt.ceo.salaries
, x = age
, geom = "histogram")
```

**Scatterplot**

*Sales vs profits*
```{r}
qplot( data = dt.ceo.salaries
, x = sales
, y = profits
, geom = "point")
```

**Barplot**

*Graduate*
```{r}
qplot( data = dt.ceo.salaries
, x = factor(grad)
, geom = "bar")
```

**Line**

*Sales vs profits*
```{r}
qplot( data = dt.ceo.salaries
, x = sales
, y = profits
, geom = "line")
```

**Facet wrap**

*Salary*
```{r}
qplot( data = dt.ceo.salaries
, x = salary
, geom = "histogram") + facet_wrap(~ grad)
```

**Customized plots**

*Salary of graduates vs non graduates*
```{r}
qplot( data = dt.ceo.salaries
, x = salary
, geom = "histogram"
, fill = factor(grad, levels = c(0,1), labels = c("Yes", "No"))) +
theme_bw() +
ylim(0,50) +
xlim(0, 4000) +
labs( title = "MY PLOT", x = "CEO Salary", y = "Number of CEOs", fill = "Grad. Degree")
```

