---
title: "Lab 12"
author: "Nikita Grabher-Meyer"
date: "11/12/2020"
output:
  word_document: default
  html_document: default
---

# Lab 12

## OLS regression on foot

*Setup*
```{r}
setwd("/Users/nikitagrabher-meyer/Desktop/PHD/Econometrics/Labs/Lab 12, Homework")

library(data.table)
library(stargazer
        )
df.OLSdat <- read.table("OLSqData.RData")
DT <- data.table(df.OLSdat) 
```

*Take a look at the data, which variables do you have?*
```{r}
summary(df.OLSdat)
summary(DT)
```

*Now use cbind to build 2 matrices: a) Using DT$const, DT$x1, DT$x2 and DT$x3 to build matrix X and b) Using DT$y5 to build matrix y*
```{r}
X <- cbind(c(DT$const), c(DT$x1), c(DT$x2), c(DT$x3))

y <- cbind(c(DT$y1))
```

*Now compute X’ (transpose of X)*
```{r}
Xt <- t(X)
```

*Next compute X’X . Report your result*
```{r}
XtX <- Xt%*%X
```

*Next compute the inverse of X’X. Report it*
```{r}
invXtX <- (XtX)^-1
```

*Next compute X’y. Report it*
```{r}
Xty <- Xt%*%y
```

*Lastly multiply the inverse of X’X with X’y. Report your result. What is this?*
```{r}
OLS <- invXtX%*%Xty
OLS
```

*Finally run a few comparisons: Regress y1 on x1, x2, and x3 in a linear model* 
```{r}
OLS1 <- lm(y1 ~ x1 + x2 + x3, data=DT)
summary(OLS1)
```

*Divide XtX by n (number of observations) and compare it to cov(DT), comment* 
```{r}
XtX/7584

covDT <- cov(DT)
covDT
```


## IV – Exercise 
```{r}
dt.IV <- read.table("IV_Data.RData")
dt.IV <- data.table(dt.IV) 
IV <- lm(y5 ~ x1 + x2, data=dt.IV)
stargazer(IV, type = "text")
```

*Looking at the Var-Cov Matrix, what bias would you expect for x1 and what bias for x2?, calculate! (Hint: the X – matrix for calculating the bias consists of x1 and x2., i.e. you can disregard the z’s and the constant for this calculation. E(X,u) is given by the “eps” - column)*
The coefficients of x1 and x2 are both positive. Also the two variable have positive correlation. The bias should be positive as well.

*Which of the X Variables is endogenous with y (e.g. a simultaneity problem)? If you can mend only one of the variables, which one would you tackle and why? You have 4 candidates that you can use as an instrumental variable for X1, but one is itself endogenous, one is a weak instrument, one is irrelevant and only one is valid. Which is which?*
1) x1 seems to be correlated with the error term, therefore it's endogenous. 2) 2) x2 also seems to be correlated with the error term, but the correlation is quite weak (0.02)
3) z1 is also endogenous (for the same reasons as above)
4) z2 is not correlated with x1, therefore it's irrelevant 
5) z3 is correlated with x1, and the covariance seems to be the highest within the different alternatives. It might be considered the best instrument
6) z4 is correlated with x1, but it is only 0.01, therefore it might be quite weak

In summary, z1, z2 and z4 should not be used as instruments, as they violate the 2 assumptions of relevance and exclusion restriction.

*Consider using Z4 as an instrument for X2: Which of the two assumptions can you test? Is it satisfied?*
We can test the relevance assumption using Pearson’s product-moment correlation test
```{r}
cor.test(dt.IV$z4, dt.IV$x1)
```

*Pick your most preferred instrument and run the 2SLS*
```{r}
first_step <- lm(x1 ~ x2 + z3, data=dt.IV)
summary(first_step)

dt.IV <- dt.IV[, x1hat:=predict(first_step, newdata=dt.IV)]

dt.IV

second_step <- lm(y5 ~ x1hat + x2, data=dt.IV)
stargazer(second_step, IV, type="text")
```

*Consider the MM-IV Estimator that we saw in class. What variables will you include in the Z matrix? Write a code that directly implements the estimator in matrix notation, using either the data.table command or the matrix command.Build a Z-matrix consisting of the instruments you want to use (including all exogeneous x); build an X-matrix consisting of the X variables; build a y-matrix that just consists of the vector y*
```{r}
z <- cbind(c(dt.IV$x2), c(dt.IV$z3))
x <- cbind(c(dt.IV$x1), c(dt.IV$x2))
y5 <- cbind(c(dt.IV$y5))

solve(t(z)%*%x)%*%(t(z)%*%y5)

library(ivreg)
IVz1 <- ivreg(y5~x1+x2|x2 + z1, data=dt.IV)
IVz2 <- ivreg(y5~x1+x2|x2 + z2, data=dt.IV)
IVz3 <- ivreg(y5~x1+x2|x2 + z3, data=dt.IV)
IVz4 <- ivreg(y5~x1+x2|x2 + z4, data=dt.IV)

stargazer (dt.IV, type="text")
```








