---
title: "Lab 6 - Homework, Part C2"
author: "Nikita Grabher-Meyer"
date: "11/2/2020"
output:
  word_document: default
  html_document: default
---

# PART C (coding)
## Part 2


### Choose at least 3 distributions and try to simulate data (e.g. 5000 draws) from a random variable that follows this distribution

*Normal distribution*
```{r}
a <- rnorm(5000)
a[1:20]
hist(a, main = "Normal Distribution")
```

*Binomial distribution: number of successes in a set of pass/fail trials with success estimated at probability p*
The function takes three arguments: number of observations you want to see, number of trials per observation, probability of success for each trial.
```{r}
b <- rbinom(5000,2,.4)
b[1:20]
```

*Poisson distribution: number of expected events for a process given we know the average rate at which events occur during a given unit of time*
The function takes three arguments: Number of observations you want to see,The estimated rate of events for the distribution; this is expressed as average events per period. 
```{r}
c <- +rpois(5000,2)
c[1:20]
```

### Draw two samples from each variable and report the sample means. Are they equal?

*Normal distribution*
```{r}
s=500 #sample size

samplea1 <- sample(a,s)
samplea2 <- sample(a,s)

mean(samplea1)
mean(samplea2)
```

*Binomial distribution*
```{r}
s=500 #sample size

sampleb1 <- sample(b,s)
sampleb2 <- sample(b,s)

mean(sampleb1)
mean(sampleb2)
```

*Poisson distribution*
```{r}
s=500 #sample size

samplec1 <- sample(c,s)
samplec2 <- sample(c,s)

mean(samplec1)
mean(samplec2)
```

The sample means (no matter which distribution is considered) are not equal.

### Follow the steps for illustrating the Central Limit Theorem using a poisson, a uniform and a normal distribution.Compare the normal, poisson and the uniform, what do you notice for 100 draws?

*Poisson distribution*
```{r}
poisson10 <- replicate (10000, {
  a <- rpois(10,5)
  mean(a)
})

poisson100 <- replicate (10000, {
  a <- rpois(100,5)
  mean(a)
})
```

*Uniform distribution*
```{r}
uniform10 <- replicate (10000, {
  b <- rpois(10,5)
  mean(b)
})

uniform100 <- replicate (10000, {
  b <- rpois(100,5)
  mean(b)
})
```

*Normal distribution*
```{r}
normal10 <- replicate (10000, {
  c <- rnorm(10)
  mean(c)
})

normal100 <- replicate (10000, {
  c <- rnorm(100)
  mean(c)
})
```

*Compare the three distributions*
```{r}
par(mfrow=c(3,2))
hist(uniform10, xlim=c(0,1), main="Uniform distribution", xlab="Sample size 10")
hist(uniform100, xlim=c(0,1), main="Uniform distribution", xlab="Sample size 100")
hist(poisson10, xlim=c(2,8), main="Poisson distribution", xlab="Sample size 10")
hist(poisson100, xlim=c(2,8), main="Poisson distribution", xlab="Sample size 100")
hist(normal10, xlim=c(-1,1), main="Normal distribution", xlab="Sample size 10")
hist(normal100, xlim=c(-1,1), main="Normal distribution", xlab="Sample size 100")
```
The central limit theorem states that if you have a population with mean μ and standard deviation σ and take sufficiently large random samples from the population, then the distribution of the sample means will be approximately normally distributed (regardless of the population distribution).


### Testing

*Load the data*
```{r}
setwd("/Users/nikitagrabher-meyer/Desktop/PHD/Econometrics/Labs/Lab 6, Homework")

library(data.table)
library(ggplot2)
library(stargazer)

load("dt_wages.RData")
dt.wages <- data.table(dt.wages)
```

*Summaries* 
```{r}
stargazer(dt.wages, type = "text")
```

*Create an estimator for the average wage by group*
```{r}
wage <- dt.wages[,list(avg_wage=mean(wage))]

wage.nowh <- dt.wages[,list(avg_wage=mean(wage)), by=nonwhite]
wage.fem <- dt.wages[,list(avg_wage=mean(wage)), by=female]

wage.nowh
wage.fem
```

*Confidence intervals*
*How to calculate the 95% confidence intervals for the population wage?*
```{r}
dt.wages[ , list (avg_wage=mean(wage) , sd_wage=sd(wage))]
```

*Create a function to calculate the 95% confidence interval*
```{r}
conf.int <- function(X){
     n <- length(X)
     error <- qt(0.975, df=n-1) * sd(X) / sqrt(n)
     mean.X <- mean(X)
     return(list(lower = mean.X - error, upper = mean.X +error))
}
```

*Apply the "conf.int" function to the wage variable*
```{r}
dt.wages[,conf.int(wage)]
```

*Confidence intervals by group*
```{r}
dt.wages[,conf.int(wage), by=nonwhite]
dt.wages[,conf.int(wage), by=female]
```

*Hypothesis testing, using the function t-test*
```{r}
dt.wages[female==1, t.test(wage,mu=5)]
```

*Are wages of men different from wages of women?*
```{r}
t.test(dt.wages[female==1,wage],dt.wages[female==0,wage])
```

*Are wages of whites different from wages of non-whites?*
```{r}
t.test(dt.wages[nonwhite==1,wage],dt.wages[nonwhite==0,wage])
```
